---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "120%"
)
```

# topR

See full documentation at [https://wuxi-nextcode.github.io/topR/](https://wuxi-nextcode.github.io/topR/)

### Installing from github using `devtools`

```{r, eval=FALSE}
#devtools::install_github("wuxi-nextcode/topR")
```

## Example
In this example we demonstrate the basic usage of the topR library.

### Load packages

First load the `topr` package, the `tidyverse` package is recommended in general, but not required for this example

```{r}
library(topR)
library(tidyverse)
```
```{r}
library(ggrepel)
```


### Loading and exploring prebuilt datasets 

Load the `gwas_CD` dataset, which is a subset of association results (SNPs with P<1e-03) for Crohn´s disease from the UK biobank.

It is highly recommended to theck the number of datapoints in your dataset before you plot, since a very large dataset will take a long time to plot.


```{r}
data(gwas_CD)
head(gwas_CD)
```

```{r}
paste("Number of SNPs in the dataset: [", length(gwas_CD$POS),"]", sep = "")
```
### Manhattan plots

Get an overview of association results for crohn's disease (CD) in a Manhattan plot

```{r fig.width=10, fig.height=4.5}
manhattan(gwas_CD)
```

### Label the top SNPs with the name of the nearest gene

Use the `get_best_snp_per_MB` function to extract the top/lead SNPs from the crohns dataset. Then include them in the manhattan call function with the variants argument.

```{r fig.width=10, fig.height=4.5}
snps_CD <- get_best_snp_per_MB(gwas_CD, thresh = 1e-09, region = 10000000)
manhattan(gwas_CD, variants=snps_CD, annotation_thresh = 1e-09)
```
## Label the position of genes of interest

```{r fig.width=10, fig.height=4.5}
genes=get_genes_by_Gene_Symbol(c("NOD2","ULK2","THADA"))
snps_CD <- get_best_snp_per_MB(gwas_CD, thresh = 1e-09, region = 10000000)
manhattan(gwas_CD, variants=snps_CD, annotation_thresh = 1e-09, highlight_genes=genes,highlight_genes_ypos = 1)
```

### Chromplot

Take a closer look at the results by chromosome. Here we plot the results on chromosome 16 only.

```{r fig.width=10, fig.height=4.5}
CHR="chr16"
chromplot(gwas_CD,chr=CHR, variants = snps_CD, annotation_thresh = 5e-09)
```

### Regionplot

Zoom in further on the chromosome plot with the `regionplot` function.

Zoom in on a gene of interest:

```{r fig.width=12, fig.height=7}
regionplot(gwas_CD, gene="NOD2")
```

Zoom in on the top hit on a chromosome

```{r fig.width=12, fig.height=7}
CHR="chr1"
top_hit=get_top_hit(gwas_CD,chr=CHR)
xmin=top_hit$POS-250000
xmax=top_hit$POS+250000
regionplot(gwas_CD, chr = CHR, xmin=xmin,xmax=xmax,show_exons = T)
```

### Display multiple datasets on the same plot

Display the output from more than one GWAS on the same plot

### Chromplot  - multiple phenotypes

Take a closer look at the results by chromosome. Here we plot the results on chromosome 16 only.

```{r fig.width=12, fig.height=4.5}
CHR="chr16"
snps_CD_IIBD <- get_best_snp_per_MB(gwas_CD_IIBD, thresh = 1e-09, region = 10000000)
chromplot(list(gwas_CD_IIBD,gwas_CD),chr=CHR, variants = list(snps_CD_IIBD,snps_CD),
          annotation_thresh = 5e-09, legend_labels=c("Crohns IIBD", "Crohns UKBB"), alpha=c(1,0.7),size=c(1.2,1))
```

### Manhattan  - multiple phenotypes

```{r fig.width=12, fig.height=6}
data(gwas_CD_IIBD)
manhattan(list(gwas_CD_IIBD,gwas_CD), legend_labels=  c("Crohns IIBD", "Crohns UKBB"), alpha = c(1, 0.5))
```

```{r fig.width=12, fig.height=6}
data(finngen_K11_CD_STRICT)
data(finngen_CHRONSMALL)
genes=get_genes_by_Gene_Symbol(c("NOD2","ULK2","THADA"))
manhattan(list(gwas_CD_IIBD,gwas_CD,finngen_K11_CD_STRICT,finngen_CHRONSMALL),
          ntop=3,
          legend_labels=  c("Crohns IIBD", "Crohns UKBB","Crohns Finngen (K11 STRICT)", "Crohns finngen (CHRONSMALL)"),
          alpha = c(1,0.5,0.3,1),
          highlight_genes = genes,
          highlight_genes_ypos = -1)
```