---
title: "Intro into the topr package "
author: "Thorhildur Juliusdottir"
output: html_notebook
---
<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Intro to the topr package}
-->

```{r, include=FALSE}
    library(topr)
    library(knitr)
    opts_chunk$set(comment=NA, fig.width=12, fig.height=5.5, message=FALSE, tidy=TRUE, dpi=170)
```
<hr>
## Example input datasets


topr comes with three example GWAS datasets, one on Ulcerative Colitis retrieved from the UKBB (<code>UC_UKBB</code>), and the other two on Crohn’s disease (<code>CD_UKBB</code> and <code>CD_FINNGEN</code>) obtained from the FinnGen and UK biobanks respectively. topr utilizes gene and exon datasets from Ensembl (GRCh38.pxx) (<code>ENSGENES</code> and <code>ENSEXONS</code>).

See <a href="https://wuxi-nextcode.github.io/topr/reference/index.html">topr reference</a> for more details on the in-built datasets.

Input datasets must include least three columns (<code>CHROM,
POS</code> and <code>P</code>), where naming of the columns is flexible
(i.e the chr label can be either chr or chrom and is case insensitive).

topr has 3 in-built datasets (GWASes), take a look at Crohn's GWAS (<code>CD_UKBB</code>) by issuing the following command:

```{r}
head(CD_UKBB)
```


The chromosome in the <code>CHROM</code> column can be represented with
or without the <i>chr</i> suffix, e.g (chr1 or 1)
<hr>

## Usage

topr's key plotting functions are <code>manhattan()</code> and <code>regionplot()</code>


### Manhattan plots

Get an overview of the Crohns association results on a Manhattan plot:


```{r manhattan, fig.height = 5.5}
manhattan(CD_UKBB)
```

Label association peaks with the nearest gene and add plot title:

```{r}
manhattan(CD_UKBB, annotate=5e-09,  title="Crohn's disease")
```


Add genes of interest to the plot

```{r}
genes=c("IL23R","NOTCH4","NOD2","JAK2","TTC33")
manhattan(CD_UKBB, annotate=5e-09, title="Crohn's disease", highlight_genes = genes)
```

Change the position of the genes of interest on the plot (the default value is 1):

```{r}
manhattan(CD_UKBB, annotate=5e-09, title="Crohn's disease", highlight_genes = genes, highlight_genes_ypos = 1.5)
```

Remove the significance threshold:

```{r}

manhattan(CD_UKBB, annotate=5e-09, title="Crohn's disease", highlight_genes = genes, highlight_genes_ypos = 1.5, sign_thresh = NULL)
```
Set multiple significance thresholds in different colors:

```{r}

manhattan(CD_UKBB, annotate=5e-09, title="Crohn's disease", highlight_genes = genes, highlight_genes_ypos = 1.5, sign_thresh =c(1e-8,1e-09,1e-10),sign_thresh_color = c("red","darkorange","magenta"))
```


Change the scale and plot colors:

```{r}
manhattan(CD_UKBB, annotate=5e-09, title="Crohn's disease", highlight_genes = genes, highlight_genes_ypos = 1.5, highlight_genes_color ="magenta", color="darkgreen", scale=1.2)
```

Change the color and angle of the labels:

```{r}
manhattan(CD_UKBB, annotate=5e-09, title="Crohn's disease", highlight_genes = genes, highlight_genes_ypos = 1.5, highlight_genes_color ="magenta", color="darkgreen", label_color = "grey40", angle=90)
```

Tidy the labels, by moving them using <code>nudge_y</code> and increase the y-axis. Increase the <code>region_size</code> for sparser gene labelling.

```{r}
manhattan(CD_UKBB, annotate=5e-09, title="Crohn's disease", highlight_genes = genes, highlight_genes_ypos = 1.5, highlight_genes_color ="magenta", color="darkgreen", label_color = "grey40", angle=90,nudge_y=2,ymax=30, region_size = 10000000)
```

Display multiple GWASes on the same plot:

```{r, fig.height = 6}
manhattan(list(UC_UKBB,CD_UKBB))
```

Set the legend labels,annotate and change plot color:
```{r, fig.height = 6}
manhattan(list(UC_UKBB,CD_UKBB), legend_labels=c("UC UKBB", "CD UKBB"), annotate=1e-12, region_size=10000000, color = c("darkblue","skyblue2"))
```

Use the <code>ntop</code> argument to control how many GWASes are shown on the top and how many at the bottom and highlight the genes of interest:

```{r, fig.height = 8}
manhattan(list(UC_UKBB,CD_UKBB), legend_labels=c("UC UKBB", "CD UKBB"), annotate=1e-12, region_size=10000000, color = c("darkblue","skyblue2"),ntop=1,
          highlight_genes = genes, highlight_genes_ypos = -0.5 , angle=90, ymax=40, ymin=-30, nudge_y = 2, label_color ="black", scale=1.1)
```

Show 3 GWASes on the same plot, using different annotation thresholds for each dataset
```{r, fig.height = 8.5}
manhattan(list(UC_UKBB,CD_UKBB,CD_FINNGEN), legend_labels=c("UC UKBB", "CD UKBB","CD FINNGEN"), annotate=c(5e-9,5e-9,1e-15), region_size=100000000, ntop=1,
          highlight_genes = genes, highlight_genes_ypos = -0.5 , angle=90, ymax=40, ymin=-30, nudge_y = 2, title="Inflammatory Bowel Disease")
```


### Region plots

Zoom-in on the region around the IL23R gene

```{r regionplot, fig.height = 8, fig.width = 12}
regionplot(CD_UKBB, gene="IL23R")
```

Annotate the top variant within each 100kb window.

```{r regionplot_annotate, fig.height = 8, fig.width = 12}
regionplot(CD_UKBB, gene="IL23R", annotate=5e-09, region_size  = 100000)
```
Annotate the top variants and draw a vertical line through their positions to further highlight their position within the genes below.

```{r regionplot_with_vline, fig.height = 8, fig.width = 12}
regionplot(CD_UKBB, gene="IL23R", annotate_with_vline = 5e-09, region_size = 100000)
```

Zoom in on the IL23R gene for multiple GWASes

```{r regionplot_multi, fig.height = 8, fig.width = 12}
regionplot(list(UC_UKBB,CD_UKBB,CD_FINNGEN), gene="IL23R", annotate_with_vline = 5e-06, legend_labels=c("UC UKBB","CD UKBB","CD FINNGEN"))
```

### Locuszoom-like plots

Locuszoom-like plot. Note that the input dataframe needs to include the R2 column with the pre-calculated r2 values, since topr does not do these calculations.

```{r locuszoom, fig.height = 8, fig.width = 12}
locuszoom(R2_CD_UKBB)
```

Annotate the variants with vlines on the plot:
```{r, fig.height = 8, fig.width = 12}
locuszoom(R2_CD_UKBB, annotate_with_vline = 1e-09, region_size = 100000)
```

## Useful functions

Extract lead/index variants from the GWAS dataset (<code>CD_UKBB</code>):

```{r}
CD_UKBB %>% get_best_snp_per_MB()
```



Annotate the lead/index variants with their nearest gene:

```{r}
CD_UKBB %>% get_best_snp_per_MB() %>%  annotate_with_nearest_gene()
```

Get genomic coordinates for a gene:

```{r}
get_gene(gene_name="IL23R")
```

Get snps within a region:

```{r}
CD_UKBB %>% get_snps_within_region(region = "chr1:67138906-67259979") %>% head(n=10)
```

Get the top variant on a chromsome:
```{r}
CD_UKBB %>% get_top_snp(chr="chr1")
```
