---
title: "Extending manhattan plots with manhattanExtra"
output: rmarkdown::html_vignette
description: |
  The topr package is a set of functions used to visualize, explore and annotate genetic association results.
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Extending manhattan plots with manhattanExtra}
  %\VignetteEncoding{UTF-8}
---

```{r ex_setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
  )
library(topr)
library(magrittr)
set.seed(0)
```

When using the regular `manhattan` function, the user can specify a P-value threshold to highlight significant associations. However, sometimes it is useful to highlight variants using both a genome-wide significant threshold and a suggestive threshold. A good use case for this could be an under powered study with a small sample size. The `manhattanExtra` extension of `manhattan` allows the user to easily do this.

There are no additional required arguments for this function, as shown below.

```{r manhextra, fig.width=12, fig.height=5}
manhattanExtra(CD_UKBB)
```

The `manhattanExtra` function can be customized using some of its optional arguments, including `genome_wide_thresh`, `suggestive_thresh`, `flank_size` & `region_size`. The `flank_size` parameter is used to denote the size on either side of a top hit that will be included in the colored grouping for that hit. The `region_size` parameter is used for gene annotation and will determine how dense the labeling is. Larger numbers will have sparser annotations.

Adjusting flank and region sizes -

```{r manhextra2, fig.width=12, fig.height=5}
manhattanExtra(
  CD_UKBB,
  flank_size = 1e6,
  region_size = 1e6
)
```

```{r manhextra3, fig.width=12, fig.height=5}
manhattanExtra(
  CD_UKBB,
  flank_size = 1e6,
  region_size = 1e8
)
```

For arguments sake, we can make the thresholds much stricter to see the results -

```{r manhextra4, fig.width=12, fig.height=5}
manhattanExtra(
  CD_UKBB,
  genome_wide_thresh = 1e-12,
  suggestive_thresh = 1e-8
)
```

```{r manhextra5, fig.width=12, fig.height=5}
manhattanExtra(
  CD_UKBB,
  genome_wide_thresh = 1e-12,
  suggestive_thresh = 1e-8,
  flank_size = 1e6,
  region_size = 1e8
)
```